numeric vs HPO

### 2022-04-22 code/r/02_B_numeric_vs_hpo_for_acute_kidney_failure.Rmd
rename files

Retrieve liver function lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_acute_kidney_failure.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_acute_kidney_failure.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FALSE)
df_labs %>% nrow()
```

@@

kidney_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

>>>>

Retrieve liver function lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_acute_kidney_failure.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_acute_kidney_failure.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FORCE_RERUN)
df_labs %>% nrow()
```

@@

kidney_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  ungroup() %>% 
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

write.csv(kidney_analysis, file = 'cache/acute_kidney_failure_analysis.csv', row.names = FALSE)

### 2022-04-22 code/r/02_C_numeric_vs_hpo_for_blood_disorder.Rmd
rename files

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
```

@@

Retrieve diagnoses information for the cohort.
```{r}
sql <- "SELECT * FROM cohort"
df_cohort <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'aplastic_anemia_diagnoses.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FALSE)

@@

Retrieve related lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_aplastic_anemia.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_aplastic_anemia.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FALSE)
df_labs %>% nrow()
```

@@

anemia_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

>>>>

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}

FORCE_RERUN <- TRUE
```

@@

Retrieve diagnoses information for the cohort.
```{r}
sql <- "SELECT * FROM cohort"
df_cohort <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'aplastic_anemia_diagnoses.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FORCE_RERUN)

@@

Retrieve related lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_aplastic_anemia.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_aplastic_anemia.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FORCE_RERUN)
df_labs %>% nrow()
```

@@

anemia_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  ungroup() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

write.csv(anemia_analysis, file = 'cache/anemia_analysis.csv', row.names = FALSE)

### 2022-04-22 code/r/02_E_numeric_vs_hpo_combined.Rmd
rename files

Load previous data analysis results
```{r}
crc_analysis <- read.csv('cache/crc_case_matched_controls_analysis.csv')
acute_kidney_failure_analysis <- read.csv('cache/acute_kidney_failure_analysis.csv')
anemia_analysis <- read.csv('cache/anemia_analysis.csv')
```

```{r}
numeric_vs_hpo <- bind_rows(crc_analysis %>% mutate(disease = 'colon cancer'), 
                            acute_kidney_failure_analysis %>% mutate(disease = 'acute kidney failure'), 
                            anemia_analysis %>% mutate(disease = 'aplastic anemia'))

@@

plot_AUC_differences_numeric_vs_hpo <- ggplot(df, aes(x = method, y = auc, group = interaction(disease, loinc))) + 
  geom_point(size = 0.5) +
  geom_line(aes(color = disease)) +
  scale_x_discrete(breaks = c('continuous', 'hpo'), labels = c('original numeric value', 'transformed to HPO')) +
  xlab("Use Lab Test Results As") +
  ylab("AUC") +
  theme_bw() + 
    theme(panel.grid = element_blank(),
          axis.text = element_text(size = 8),
          axis.title = element_text(size = 8), 
          strip.text.x = element_text(margin = margin(0.3, 0, 0.3, 0), size = 5),
          plot.title = element_text(size = 8))

ggsave(filename = 'cache/plot_AUC_differences_numeric_vs_hpo.pdf', width = 5, height = 3)

>>>>

Load previous data analysis results
```{r}
liver_analysis <- read.csv('cache/liver_analysis.csv')
crc_analysis <- read.csv('cache/crc_case_matched_controls_analysis.csv')
acute_kidney_failure_analysis <- read.csv('cache/acute_kidney_failure_analysis.csv')
anemia_analysis <- read.csv('cache/anemia_analysis.csv')
```

```{r}
numeric_vs_hpo <- bind_rows(liver_analysis %>% mutate(disease = 'abnormal liver function'),
                            crc_analysis %>% mutate(disease = 'colon cancer'), 
                            acute_kidney_failure_analysis %>% mutate(disease = 'acute kidney failure'), 
                            anemia_analysis %>% mutate(disease = 'aplastic anemia'))

@@

plot_AUC_differences_numeric_vs_hpo <- ggplot(df, aes(x = method, y = auc, group = interaction(disease, loinc))) + 
  geom_point(size = 1) +
  geom_line(aes(color = disease)) +
  scale_x_discrete(breaks = c('continuous', 'hpo'), labels = c('original numeric value', 'transformed to HPO')) +
  xlab("Use Lab Test Results As") +
  ylab("AUC") +
  theme_bw() + 
    theme(panel.grid = element_blank(), legend.position = 'none',
          axis.text = element_text(size = 8),
          axis.title = element_text(size = 8), 
          strip.text.x = element_text(margin = margin(0.3, 0, 0.3, 0), size = 5),
          plot.title = element_text(size = 8)) +
  facet_wrap(~ disease)

ggsave(filename = 'cache/plot_AUC_differences_numeric_vs_hpo.pdf', width = 6, height = 6)
